---
title: "AUMGC for *Colletotrichum fioriniae* and *Colletotrichum karstii*"
author: "Paulo Lichtemberg & Walmes Zeviani"
date: "31 de Janeiro de 2020"
output:
  html_document:
    theme: cosmo
    highlight: haddock
    toc: true
---

# Session definitions

```{r, message = FALSE, results = "hide"}
# List of packages.
pkg <- c("tidyverse", "agricolae", "emmeans", "segmented")

# Loads each package.
sapply(pkg, FUN = library, character.only = TRUE, logical.return = TRUE)
```

```{r, include = FALSE}
devtools::load_all()
```

# Area under the micelial growth curve

## Exploratory data analysis

```{r}
# Prepare table of data.
ogt <- as_tibble(RDASC::pistachio_anthracnose$ogrotem)
ogt$exp <- factor(ogt$exp)
attr(ogt, "spec") <- NULL
str(ogt)

# Get the mean diameter.
ogt$diameter <- (ogt$mm1 + ogt$mm2)/2

# Curves for each isolate according to experimental conditions.
ggplot(data = ogt,
       mapping = aes(x = day,
                     y = diameter,
                     color = iso)) +
    facet_grid(facets = exp ~ tem) +
    geom_point() +
    stat_summary(geom = "line", fun.y = "mean") +
    labs(x = "Days",
         y = "Diameter (mm)",
         color = "Isolate")

# Loads a function to calculate area under a curve by trapezoidal
# method.
source("https://raw.githubusercontent.com/walmes/wzRfun/master/R/auc.R")
args(auc)
body(auc)

# Frequency tables.
xtabs(~iso + spp, data = ogt)
ftable(xtabs(~exp + tem + iso, data = ogt))

# AUC by each experimental unit = exp > spp > iso > tem > rep.
aumgc <- ogt %>%
    group_by(exp, spp, iso, tem, rep) %>%
    summarise(auc = auc(time = day, resp = diameter)) %>%
    ungroup()
aumgc

# Convert come experimental controled variables to factor.
aumgc <- aumgc %>%
    mutate_at(vars(exp, spp, iso, rep), factor)
str(aumgc)

# Visualize.
ggplot(data = aumgc,
       mapping = aes(x = tem, y = auc, color = exp)) +
    facet_wrap(facets = ~iso) +
    geom_point() +
    stat_summary(geom = "line", fun.y = "mean") +
    labs(x = "Temperature",
         y = "AUC",
         color = "Experiment")

# Removes outlier.
aumgc[with(aumgc,
           iso == "12D46" &
           exp == 3 &
           tem == 25 &
           auc < 50), ]$auc <- NA
```

## Optimal temperature for each isolate

  * `exp`: represents the contour conditions of each experiment, so it
    represents the blocking factor.
  * `iso`: is the effect of isolate.
  * `rep`: is each independent replication of experiment $\times$
    isolate $\times$ temperature. Response could be averaged across
    replications.

```{r}
# Uses a "naive" model (in terms of variance components) to check the
# model assumptions.

# Averages the AUC.
aumgcm <- aumgc %>%
    group_by(exp, iso, tem) %>%
    summarise(aucm = mean(auc, na.rm = TRUE))

# Visualize.
ggplot(data = aumgcm,
       mapping = aes(x = tem, y = aucm, color = exp)) +
    facet_wrap(facets = ~iso) +
    geom_point() +
    geom_line() +
    labs(x = "Temperature",
         y = "AUC",
         color = "Experiment")

# Model fitting.
m0 <- lm(aucm ~ exp + iso * factor(tem),
         data = aumgcm)

# Residual plots.
par(mfrow = c(2, 2))
plot(m0)
layout(1)

# NOTE: no strong evidence against model assumptions.
```

## Optimal temperature estimation

Estimation of optimal temperature is done for each experimental
condition: experiment $\times$ isolate.  After estimation, a model for
hyphotesis on equality of the optimal values will be employed.

```{r}
# Each experimental cell.
gg <- ggplot(data = aumgcm,
       mapping = aes(x = tem, y = aucm)) +
    facet_grid(facets = exp ~ iso) +
    geom_point() +
    expand_limits(y = c(0, NA)) +
    geom_hline(yintercept = 25, linetype = 3) +
    labs(x = "Temperature",
         y = "AUC")

gg +
    geom_smooth(method = "lm",
                formula = y ~ poly(x, degree = 4),
                se = FALSE)


# ATTENTION FIXME: the values of diameter must be subtracted from the
# plug initial diameter? Is 25 mm?
```

A 4 degree polynomial will be used to fit the (mean) AUC as a function
of temperature.  After fitting, the optimal temperature will be
determined by a numerical optimation method.

```{r}
#-----------------------------------------------------------------------
# Polynomial approach.

# Function that fits a model equation to data and determine the optimal
# temperature.
fit_poly <- function(data,
                         formula = aucm ~ poly(tem, degree = 4),
                         interval = c(10, 35),
                         maximum = TRUE) {
    m0 <- lm(formula = formula, data = data)
    pred <- function(x) predict(m0, newdata = list(tem = x))
    t_opt <- optimise(f = pred,
                      interval = interval,
                      maximum = maximum)
    return(data.frame(t_opt = t_opt$maximum,
                      sqr = deviance(m0),
                      ll = logLik(m0)))
}

# # Testing.
# fit_poly(data = aumgcm[with(aumgcm, iso == "11J23" & exp == "1"),
#                        c("aucm", "tem" )])

#-----------------------------------------------------------------------
# Bilinear approach.

fit_bilinear <- function(data,
                         start = list(b0 = -50,
                                      b1 = 8,
                                      xk = 26,
                                      b2 = -17)) {
    # Uses `nls()` function.
    n0 <- nls(aucm ~ b0 +
                  (b1 * tem) * (tem <= xk) +
                  (b1 * xk + b2 * (tem - xk)) * (tem > xk),
              start = start,
              data = data)
    return(data.frame(t_opt = coef(n0)["xk"],
                      sqr = deviance(n0),
                      ll = logLik(n0)))
}

# fit_tem_bilinear <- function(data) {
#     # Uses `segmented` package.
#     m0 <- lm(formula = aucm ~ tem, data = data)
#     s0 <- segmented(m0, seg.Z = ~tem, psi = 26)
#     return(data.frame(t_opt = s0$psi[1, "Est."],
#                       sqr = deviance(s0),
#                       ll = logLik(s0)))
# }

# # Testing.
# fit_bilinear(data = aumgcm[with(aumgcm, iso == "11J23" & exp == "2"), ])

#-----------------------------------------------------------------------
# Fits for each experimental cell.

temps <- aumgcm %>%
    group_by(iso, exp) %>%
    summarise(topt_p = list(
                  fit_poly(data = data.frame(aucm, tem))),
              topt_b = list(
                  fit_bilinear(data = data.frame(aucm, tem)))) %>%
    ungroup()

temps <- temps %>%
    gather(key = "approach", value = "measures", topt_p, topt_b) %>%
    unnest()

# Which approach has the smallest deviance (best fit)?
ggplot(temps,
       mapping = aes(x = approach,
                     y = sqr,
                     color = exp,
                     group = exp)) +
    facet_wrap(facets = ~iso) +
    geom_point() +
    geom_line()

# What are the optimal temperature values by each approach?
ggplot(temps,
       mapping = aes(x = approach,
                     y = t_opt,
                     color = exp,
                     group = exp)) +
    facet_wrap(facets = ~iso) +
    geom_point() +
    geom_line()

# See the results for the polynomial approach.
gg +
    geom_smooth(method = "lm",
                formula = y ~ poly(x, degree = 4),
                se = FALSE,
                color = "purple",
                size = 0.6) +
    geom_vline(data = filter(temps, approach == "topt_p"),
               mapping = aes(xintercept = t_opt),
               color = "purple",
               linetype = 2)

# To configure `nls()` call inside `geom_smooth()`.
nls_args <- list(formula = y ~ b0 +
                     (b1 * x) * (x <= xk) +
                     (b1 * xk + b2 * (x - xk)) * (x > xk),
                 start = list(b0 = -50,
                              b1 = 8.5,
                              xk = 26,
                              b2 = -16))

gg +
    geom_smooth(method = "nls",
                method.args = nls_args,
                se = FALSE,
                color = "purple",
                size = 0.6) +
    geom_vline(data = filter(temps, approach == "topt_b"),
               mapping = aes(xintercept = t_opt),
               color = "purple",
               linetype = 2)
```

## Isolate comparison

```{r}
# Filter by the chosen approach.
da <- temps %>%
    filter(approach == "topt_p") %>%
    select(iso, exp, t_opt)

# Fit the model according to the experiment layout.
m0 <- lm(t_opt ~ exp + iso, data = da)

# Residual plots.
par(mfrow = c(2, 2))
plot(m0)
layout(1)

# ANOVA table.
anova(m0)

# Tukey HSD test on isolate means.
tk <- HSD.test(m0, trt = "iso", group = TRUE, console = FALSE)
tk <- tk$groups %>%
    rownames_to_column() %>%
    rename("iso" = "rowname")
tk

# Least squares means.
lsm <- emmeans(m0, specs = "iso")
lsm <- print(lsm) %>%
    mutate(iso = as.character(iso))
str(lsm)

# Merge these two tables.
tb_means <- full_join(x = lsm, y = tk, by = "iso")
tb_means
```

```{r, fig.height = 3.5}
# Segment plot displaying means with confidence interval.
ggplot(data = tb_means,
       mapping = aes(y = fct_reorder(iso, emmean), x = emmean)) +
    geom_point() +
    geom_errorbarh(mapping = aes(xmin = lower.CL, xmax = upper.CL),
                  height = 0.1) +
    geom_text(mapping = aes(label = sprintf("%0.1f%s", emmean, groups)),
              vjust = 0, nudge_y = 0.10) +
    labs(y = "Isolate",
         x = expression("Temperature" ~ (degree * C)))
```

# Sporulation capability

## Exploratory data analysis

```{r}
# Prepare table of data.
spo <- as_tibble(RDASC::pistachio_anthracnose$spo_vt)
spo$exp <- factor(spo$exp)
attr(spo, "spec") <- NULL
str(spo)

# Summaries.
summary(spo)

# Table frequency.
xtabs(~spp + iso, data = spo)
xtabs(~iso + cv, data = spo)

# Creating the context or offset variable for the count (spo).
spo$offset <- with(spo, slice *
                        c(16, 4)[match(index, c(160000, 10000))])

# Creates the variable concentration.
# Numbers correspond to 1000 in slide and 16 in cells.
spo$con <- with(spo, 1000 * 16 * spo/offset)

ggplot(data = spo,
       mapping = aes(x = tem, y = con, color = iso)) +
    facet_grid(facets = exp ~ cv) +
    geom_jitter(width = 0.5) +
    stat_summary(geom = "line", fun.y = "mean") +
    scale_y_log10()
```

## Model fitting

Data were transform to log scale to avoid violations in model
assumption.

```{r}
# Convert from numerical to factor.
spo$tem <- factor(spo$tem)

# Fit the model.
# m0 <- lm(con ~ exp + iso * cv * tem, data = spo)
# MASS::boxcox(m0)
m0 <- lm(log10(con) ~ exp + iso * cv * tem, data = spo)

# Check for any violation.
par(mfrow = c(2, 2))
plot(m0)
layout(1)

# Anova table.
anova(m0)

# Update model dropping no relevant experimental terms.
m1 <- update(m0, . ~ exp + iso)
anova(m1, m0)

# Gets the estimated marginal means for isolates.
emm <- emmeans(m1, specs = ~iso)
emm %>% as.data.frame()

# Values in the response scale.
tb_means <- multcomp::cld(emm) %>%
    as.data.frame() %>%
    mutate_at(c("emmean", "lower.CL", "upper.CL"), exp)
tb_means$cld <- c("b", "a")
tb_means
```

```{r, fig.height = 2}
# Results.
ggplot(data = tb_means,
       mapping = aes(y = iso, x = emmean)) +
    geom_point() +
    geom_errorbarh(mapping = aes(xmin = lower.CL, xmax = upper.CL),
                  height = 0.1) +
    geom_text(mapping = aes(label = sprintf("%0.2f%s", emmean, cld)),
              vjust = 0,
              nudge_y = 0.05,
              show.legend = FALSE) +
    labs(x = "Number of spores",
         # x = expression(Log[10] ~ "of sporulation"),
         y = "Isolates")
```

# Conidia characterization

## Exploratory data analysis

```{r}
cs <- as_tibble(RDASC::pistachio_anthracnose$cs)
cs$exp <- factor(cs$exp)
attr(cs, "spec") <- NULL
str(cs)

summary(cs)

# Curves for each isolate according to experimental conditions.
ggplot(data = cs,
       mapping = aes(x = len,
                     y = wid,
                     color = kare)) +
    facet_wrap(facets = ~exp) +
    geom_point()

# Find the convex hull of the points being plotted.
cs_hull <- cs %>%
    group_by(exp, kare) %>%
    slice(chull(len, wid))

ggplot(data = cs,
       mapping = aes(x = len,
                     y = wid,
                     color = kare)) +
    facet_wrap(facets = ~exp) +
    geom_polygon(data = cs_hull, alpha = 0.1) +
    geom_point() +
    labs(x = expression("Conidia length" ~ (mu * m)),
         y = expression("Conidia width" ~ (mu * m)))

ggplot(data = cs,
       mapping = aes(x = len,
                     color = kare)) +
    facet_wrap(facets = ~exp) +
    geom_density()

ggplot(data = cs,
       mapping = aes(x = wid,
                     color = kare)) +
    facet_wrap(facets = ~exp) +
    geom_density()

ggplot(data = cs,
       mapping = aes(x = vol,
                     color = kare)) +
    facet_wrap(facets = ~exp) +
    geom_density()

cs %>%
    gather(key = "variable", value = "value", len, wid, vol) %>%
    group_by(variable) %>%
    mutate(value = scale(value)) %>%
    ggplot(mapping = aes(x = value,
                         color = kare)) +
    facet_grid(facets = exp ~ variable, scales = "free") +
    geom_density() +
    geom_rug()

cs %>%
    gather(key = "variable", value = "value", len, wid, vol) %>%
    group_by(variable) %>%
    mutate(value = scale(value)) %>%
    ggplot(mapping = aes(x = value,
                         color = kare)) +
    facet_grid(facets = exp ~ variable, scales = "free") +
    stat_ecdf() +
    geom_rug()
```

## Conidia length

```{r}
xtabs(~kare + exp, data = cs)

m0 <- lm(len ~ exp * kare, data = cs)

par(mfrow = c(2, 2))
plot(m0)
layout(1)

anova(m0)

# DANGER: Tukey test is based on geometric mean of the number of
# replicates. Since the number of replications are very enequal, this
# aproximation is not suitable.
tt <- HSD.test(m0, "kare", console = FALSE)
tt$groups

emm <- emmeans(m0, specs = ~kare)
tb_means <- multcomp::cld(emm) %>%
    as.data.frame()
tb_means

tb_means$cld <- c("a", "b")[c(2, 2, 2, 2, 2, 2, 1)]

# Segment plot displaying means with confidence interval.
ggplot(data = tb_means,
       mapping = aes(y = fct_reorder(kare, emmean), x = emmean)) +
    geom_point() +
    geom_errorbarh(mapping = aes(xmin = lower.CL, xmax = upper.CL),
                  height = 0.1) +
    geom_text(mapping = aes(label = sprintf("%0.2f%s", emmean, cld)),
              vjust = 0, nudge_y = 0.05) +
    labs(x = expression("Conidia length" ~ (mu * m)),
         y = "Isolates")
```

## Conidia width

```{r}
m0 <- lm(wid ~ exp * kare, data = cs)

par(mfrow = c(2, 2))
plot(m0)
layout(1)

anova(m0)

# DANGER: Tukey test is based on geometric mean of the number of
# replicates. Since the number of replications are veru enequal, this
# aproximation is not suitable.
# tt <- HSD.test(m0, "kare", console = TRUE)
# tt$groups

emm <- emmeans(m0, specs = ~kare | exp)
emm

m1 <- aov(wid ~ exp + kare + Error(exp/kare), data = cs)
summary(m1)

emmeans(m1, specs = ~kare)


grid <- attr(emm, "grid")
L <- attr(emm, "linfct")
rownames(L) <- grid$kare

tb_means <- rbind(wzRfun::apmc(L[grid$exp == "1", ], m0, focus = "kare"),
                  wzRfun::apmc(L[grid$exp == "2", ], m0, focus = "kare"))

tb_means <- cbind(rbind(filter(grid, exp == "1"),
                        filter(grid, exp == "2")),
                  select(tb_means, -kare))
tb_means

# Segment plot displaying means with confidence interval.
ggplot(data = tb_means,
       mapping = aes(y = fct_reorder(kare, fit), x = fit)) +
    facet_wrap(facets = ~exp) +
    geom_point() +
    geom_errorbarh(mapping = aes(xmin = lwr, xmax = upr),
                  height = 0.1) +
    geom_text(mapping = aes(label = sprintf("%0.2f%s", fit, cld)),
              vjust = 0, nudge_y = 0.05) +
    labs(x = expression("Conidia width" ~ (mu * m)),
         y = "Isolates")
```

## Conidia volume

```{r}
m0 <- lm(vol^(1/3) ~ exp * kare, data = cs)

# MASS::boxcox(m0)

par(mfrow = c(2, 2))
plot(m0)
layout(1)

anova(m0)

# DANGER: Tukey test is based on geometric mean of the number of
# replicates. Since the number of replications are veru enequal, this
# aproximation is not suitable.
# tt <- HSD.test(m0, "kare", console = TRUE)
# tt$groups

emm <- emmeans(m0, specs = ~kare + exp)
grid <- attr(emm, "grid")
L <- attr(emm, "linfct")
rownames(L) <- grid$kare

tb_means <- rbind(wzRfun::apmc(L[grid$exp == "1", ], m0, focus = "kare"),
                  wzRfun::apmc(L[grid$exp == "2", ], m0, focus = "kare"))

tb_means <- cbind(rbind(filter(grid, exp == "1"),
                        filter(grid, exp == "2")),
                  select(tb_means, -kare))
tb_means

# Segment plot displaying means with confidence interval.
ggplot(data = tb_means,
       mapping = aes(y = fct_reorder(kare, fit), x = fit)) +
    facet_wrap(facets = ~exp) +
    geom_point() +
    geom_errorbarh(mapping = aes(xmin = lwr, xmax = upr),
                  height = 0.1) +
    geom_text(mapping = aes(label = sprintf("%0.2f%s", fit, cld)),
              vjust = 0, nudge_y = 0.05) +
    labs(x = expression("Cubic root of conidia volume" ~ (mu * m)),
         y = "Isolates")
```

## Principal components (bonus)

```{r, eval = FALSE}
library(factoextra)

fit <- prcomp(cs[, c("len", "wid")], scale = TRUE)

fviz_pca_biplot(fit,
                col.ind = cs$kare,
                col.var = "black",
                addEllipses = TRUE,
                ellipse.type = "confidence",
                ellipse.level = 0.95)
```

# Germination

## Exploratory data analysis

```{r}
germ <- as_tibble(RDASC::pistachio_anthracnose$ogertem)
germ$exp <- factor(germ$exp)
germ$tim <- factor(germ$tim)
attr(germ, "spec") <- NULL
str(germ)

summary(germ)

# Curves for each isolate according to experimental conditions.
ggplot(data = germ,
       mapping = aes(x = tem,
                     y = ger/of,
                     color = tim)) +
    facet_grid(facets = exp ~ iso) +
    geom_point() +
    stat_summary(geom = "line", fun.y = "mean")

ggplot(data = filter(germ, tem > 10),
       mapping = aes(x = tem,
                     y = ger/of,
                     color = tim)) +
    facet_grid(facets = exp ~ iso) +
    geom_jitter(width = 1, height = 0) +
    # stat_summary(geom = "line", fun.y = "mean") +
    geom_smooth(method = "glm",
                formula = y ~ poly(x, degree = 3),
                # formula = y ~ splines::bs(x, df = 2, degree = 3),
                # formula = y ~ splines::ns(x, df = 3),
                se = FALSE,
                method.args = list(family = "quasibinomial"))
```

## Fit curves for each experimental point

```{r}
# Function that fits a model equation to data and determine the optimal
# temperature.

fit_tem_poly <- function(data) {
    m0 <- glm(formula = cbind(ger, of - ger) ~ poly(tem, degree = 3),
              data = data,
              family = quasibinomial)
    pval <- tail(summary(m0)$coeff[, "Pr(>|t|)"], n = 1)
    if (pval > 0.05) {
        update(m0, . ~ poly(tem, degree = 2))
    }
    pred <- function(x) predict(m0,
                                newdata = list(tem = x),
                                type = "response")
    t_opt <- optimise(f = pred,
                      interval = c(10, 35),
                      maximum = TRUE)
    return(data.frame(temp_opt = t_opt$maximum,
                      germ_opt = t_opt$objective,
                      n_par = length(coef(m0)),
                      deviance = deviance(m0)))
}

data = germ[with(germ, iso == "11J23" & exp == "1" & tim == "6" & tem > 10), ]

# Testing.
fit_tem_poly(data = germ[with(germ, iso == "11J23" & exp == "1" & tim == "6" & tem > 10), ])
#

temp_opt <- germ %>%
    filter(tem > 10 & !(tim == "12" & exp == "2" & iso == "3G23")) %>%
    group_by(exp, iso, tim) %>%
    do({
        fit_tem_poly(.)
    }) %>%
    as.data.frame()

ggplot(data = temp_opt,
       mapping = aes(x = temp_opt, y = germ_opt, color = tim)) +
    facet_wrap(facets = ~iso) +
    geom_point()

ggplot(data = temp_opt,
       mapping = aes(x = iso, y = temp_opt, color = tim, group = tim)) +
    geom_point() +
    stat_summary(geom = "line", fun.y = "mean") +
    labs(y = "Optimal temperature for germination",
         x = "Isolates")

ggplot(data = temp_opt,
       mapping = aes(x = iso, y = germ_opt, color = tim, group = tim)) +
    geom_point() +
    stat_summary(geom = "line", fun.y = "mean") +
    labs(y = "Maximum germination",
         x = "Isolates")
```

## Isolate comparison

```{r}
m0 <- lm(temp_opt ~ exp + iso * tim, data = temp_opt)

par(mfrow = c(2, 2))
plot(m0)
layout(1)

anova(m0)

emm <- emmeans(m0, specs = ~iso)
L <- attr(emm, "linfct")
rownames(L) <- levels(germ$iso)

tb_means <- wzRfun::apmc(L, m0, focus = "iso")
tb_means

# Segment plot displaying means with confidence interval.
ggplot(data = tb_means,
       mapping = aes(y = fct_reorder(iso, fit), x = fit)) +
    geom_point() +
    geom_errorbarh(mapping = aes(xmin = lwr, xmax = upr),
                  height = 0.1) +
    geom_text(mapping = aes(label = sprintf("%0.2f%s", fit, cld)),
              vjust = 0,
              nudge_y = 0.05,
              show.legend = FALSE) +
    labs(x = expression("Optimal temperature for germination" ~ (""^degree * C)),
         y = "Isolates")

# Analysis on the logit scale of the estimated germination.
temp_opt$logit_germ_opt <- binomial()$linkfun(temp_opt$germ_opt)
m0 <- lm(logit_germ_opt ~ exp + iso * tim, data = temp_opt[-8, ])

# m0 <- lm(germ_opt ~ exp + iso, data = temp_opt)

# MASS::boxcox(m0)

par(mfrow = c(2, 2))
plot(m0)
layout(1)

anova(m0)

m0 <- lm(logit_germ_opt ~ exp + iso + tim, data = temp_opt[-8, ])

emm <- emmeans(m0, specs = ~tim)
L <- attr(emm, "linfct")
rownames(L) <- levels(germ$tim)

tb_means <- wzRfun::apmc(L, m0, focus = "tim") %>%
    mutate_at(c("fit", "lwr", "upr"), binomial()$linkinv)
tb_means

emm <- emmeans(m0, specs = ~iso)
L <- attr(emm, "linfct")
rownames(L) <- levels(germ$iso)

tb_means <- wzRfun::apmc(L, m0, focus = "iso") %>%
    mutate_at(c("fit", "lwr", "upr"), binomial()$linkinv)
tb_means

# ATTENTION: these values of germination is a mean of the predicted
# germination for the levels of tim. So, interpret them with caution. In
# fact, what is relevant is the contrast between isolates that represent
# the isolate effect.

# Segment plot displaying means with confidence interval.
ggplot(data = tb_means,
       mapping = aes(y = fct_reorder(iso, fit), x = fit)) +
    geom_point() +
    geom_errorbarh(mapping = aes(xmin = lwr, xmax = upr),
                  height = 0.1) +
    geom_text(mapping = aes(label = sprintf("%0.2f%s", fit, cld)),
              vjust = 0,
              nudge_y = 0.05,
              show.legend = FALSE) +
    labs(x = expression("Maximum germination"),
         y = "Isolates")

# Get the means for all combinations despite the effects are additive.
emm <- emmeans(m0, specs = ~iso + tim) %>%
    as.data.frame() %>%
    rename(fit = "emmean", lwr = "lower.CL", upr = "upper.CL") %>%
    mutate_at(c("fit", "lwr", "upr"), binomial()$linkinv)

pd <- position_dodge(0.1)
ggplot(data = emm,
       mapping = aes(x = fct_reorder(iso, fit),
                     y = fit,
                     color = tim,
                     group = tim)) +
    geom_point(position = pd) +
    geom_errorbar(mapping = aes(ymin = lwr, ymax = upr),
                  position = pd,
                  width = 0.1) +
    geom_text(mapping = aes(label = sprintf("%0.2f", fit)),
              position = pd,
              hjust = -0.2,
              show.legend = FALSE) +
    labs(y = expression("Maximum germination"),
         x = "Isolates",
         color = "Incubation\nperiod (h)")
```

# Apressory formation

# Patogenicity in vivo

# Time of sensibility

# Weather data log
