---
title: >
  Effect of fungicide sprays programs and pistachio hedging on fruit
  stain levels caused by *Alternaria* late blight in commercial
  pistachio orchard of Tulare County, California.
author: >
  [Paulo S. F. Lichtemberg](http://lattes.cnpq.br/8132272273348880)</br>
  Ryan D. Puckett</br>
  [Walmes M. Zeviani](http://www.leg.ufpr.br/doku.php/pessoais:walmes)</br>
  Connor G. Cunningham</br>
  Themis J. Michailides
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Effect of fungicide sprays programs and pistachio hedging on fruit stain levels caused by \emph{Alternaria} late blight in commercial pistachio orchard of Tulare County, California.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Session Definition

```{r, message = FALSE, results = "hide"}
# https://github.com/walmes/wzRfun
# devtools::install_github("walmes/wzRfun")

library(lattice)
library(latticeExtra)
library(plyr)
library(reshape)
library(doBy)
library(multcomp)
library(lme4)
library(lmerTest)
library(wzRfun)
```
```{r, eval = FALSE}
library(wzCoop)
```
```{r setup, include = FALSE}
# Setting to english.
source("config/setup.R")
tbn_ <- captioner(prefix = "Table")
fgn_ <- captioner(prefix = "Figure")
tbl_ <- function(label) tbn_(label, display = "cite")
fgl_ <- function(label) fgn_(label, display = "cite")
```

# Exploratory data analysis

```{r}
# Data strucuture.
str(quality_ake_b)

# Short names are easier to use.
da <- quality_ake_b
da$yr <- factor(da$yr)

intlev <- c(c0 = "[0]",
            c1 = "[1, 10]",
            c2 = "[11, 35]",
            c3 = "[36, 64]",
            c4 = "[65, 100]")

# Creating the blocks.
da$block <- with(da, {
    factor(ifelse(as.integer(gsub("\\D", "", plo)) > 2, "II", "I"))
})

# Creating unique levels for trees.
da$tre <- with(da,
               interaction(hed, tra, block, tre, drop = TRUE))

# Stack data to plot variables.
db <- melt(da,
           id.vars = 1:6,
           measure.vars = grep("c\\d", names(quality_ake_b)),
           na.rm = TRUE)
names(db)[ncol(db) - 1:0] <- c("categ", "freq")

useOuterStrips(
    xyplot(freq ~ categ | hed + yr,
           groups = tra,
           data = db,
           xlab = "Percentage of the shell surface discolored",
           ylab = "Number of fruits in each class (n = 100)",
           jitter.x = TRUE,
           scales = list(x = list(labels = intlev,
                                  alternating = 1)),
           auto.key = list(columns = 2,
                           title = "Fungicide",
                           cex.title = 1.1),
           type = c("p", "a")))

useOuterStrips(
    xyplot(freq ~ categ | tra + yr,
           groups = hed,
           data = db,
           xlab = "Percentage of the shell surface discolored",
           ylab = "Number of fruits in each class (n = 100)",
           jitter.x = TRUE,
           scales = list(x = list(labels = intlev,
                                  rot = 45,
                                  alternating = 1)),
           auto.key = list(columns = 2,
                           title = "Hed",
                           cex.title = 1.1),
           type = c("p", "a")))

i <- grep("^c\\d$", x = names(da))

# useOuterStrips(
#     splom(~da[, i] | hed + yr,
#           groups = tra,
#           type = c("p", "r"),
#           data = da))
```

```{r}
# Acumulated proportion.
cml <- t(apply(da[, i], MARGIN = 1, FUN = cumsum))
cml <- cml[, -length(i)]
colnames(cml) <- paste0("p", 1:ncol(cml))

# Add acummulated proportions.
da <- cbind(da, as.data.frame(cml))

db <- melt(da,
           id.vars = 1:6,
           measure.vars = colnames(cml),
           na.rm = TRUE)
names(db)[ncol(db) - 1:0] <- c("categ", "prop")
str(db)

accumlev <- c(p1 = "0",
              p2 = "<=10",
              p3 = "<=35",
              p4 = "<=65")

useOuterStrips(
    xyplot(prop ~ categ | hed + yr,
           groups = tra,
           data = db,
           xlab = "Percentage of the shell surface discolored",
           ylab = "Number of fruits in each class (n = 100)",
           jitter.x = TRUE,
           auto.key = list(columns = 2,
                           title = "Fungicide",
                           cex.title = 1.1),
           scales = list(x = list(labels = accumlev,
                                  alternating = 1)),
           type = c("p", "a")))

useOuterStrips(
    xyplot(prop ~ categ | tra + yr,
           groups = hed,
           data = db,
           xlab = "Percentage of the shell surface discolored",
           ylab = "Number of fruits in each category (n = 100)",
           jitter.x = TRUE,
           auto.key = list(columns = 2,
                           title = "Hed",
                           cex.title = 1.1),
           scales = list(x = list(labels = accumlev,
                                  alternating = 1)),
           type = c("p", "a")))
```

# Analysis of the accumulated proportion

The plots above showed that the greatest difference among levels of
treaments, either hed or fungicides, occurs for the accumulated
proportions at class $\leq 35$ (`p2`).  Because of this, the analysis
will be for this response variable.

The total number of independent observations results from the calculus
$$
\begin{aligned}
  288 \text{ observations} &= 2 \text{ years }
   \times 2 \text{ blocks } \times \\
   &\quad\,\, 2 \text{ heds } \times 4 \text{ fungicides } \times \\
   &\quad\,\, 3 \text{ trees } \times 3 \text{ bags}.\\
\end{aligned}
$$

```{r echo = -c(1:2), fig.cap = cap}
cap <-
    "Number of fruits with 35% or less of the shell surface discolored as a function of the fungicide and hed level for two years."
cap <- fgn_("num35less", cap)

da$y <- da$p2
da <- da[!is.na(da$y), ]

xyplot(p2 ~ tra | yr,
       groups = hed,
       data = da,
       type = c("p", "a"),
       xlab = "Fungicides",
       ylab = expression("Number of fruits with" <= "35% of the shell surface discolored"),
       jitter.x = TRUE,
       auto.key = list(columns = 2,
                       title = "Hed",
                       cex.title = 1.1))

# xyplot(p2 ~ yr | tre,
#        data = da,
#        xlab = "Years",
#        ylab = "Number of fruits with 35% or less damage",
#        as.table = TRUE,
#        jitter.x = TRUE)
```

The analysis are being separated for each year. First, although the
choosen response variable is limited or bounded (because it is a sample
proportion), it doesn't show problematic departures from a linear
(gaussian) model.  Linear models are more flexible for analysing
clutered data as such.

```{r}
da15 <- subset(da, yr == "2015")
da16 <- subset(da, yr == "2016")

# Quasi-binomial model.
# m0 <- glm(cbind(p2, 100 - p2) ~ block + yr * hed * tra,
#           family = quasibinomial,
#           data = da)

# Just to check if the model assumptions are meet.
m0 <- lm(p2 ~ block + yr * hed * tra,
         data = da)
par(mfrow = c(2, 2))
plot(m0); layout(1)

# Normal approximation goes well.
```

A simple linear models is fitted only to check how the residuals behave.
No evidence of departures from models assumptions were found.

```{r}
# Year 2015.
m15 <- lmer(p2 ~ block + hed * tra + (1 | tre),
           data = da15)

# r <- residuals(m15)
# f <- fitted(m15)
# plot(r ~ f)
# qqnorm(r)

anova(m15)
summary(m15)

# Year 2016.
m16 <- lmer(p2 ~ block + hed * tra + (1 | tre),
           data = da16)

# r <- residuals(m16)
# f <- fitted(m16)
# plot(r ~ f)
# qqnorm(r)

anova(m16)
summary(m16)
```

For both years, no effect were found for the experimental factors.

```{r}
# Represent the results with interval confidence on ls means.
cmp <- list()

lsm <- LSmeans(m15, effect = "hed")
grid <- equallevels(lsm$grid, da)

L <- lsm$K
rownames(L) <- grid[, 1]

cmp$hed <-
    rbind(
        cbind(yr = "2015", apmc(L, model = m15, focus = "hed")),
        cbind(yr = "2016", apmc(L, model = m16, focus = "hed")))

lsm <- LSmeans(m15, effect = "tra")
grid <- equallevels(lsm$grid, da)

L <- lsm$K
rownames(L) <- grid[, 1]

cmp$tra <-
    rbind(
        cbind(yr = "2015", apmc(L, model = m15, focus = "tra")),
        cbind(yr = "2016", apmc(L, model = m16, focus = "tra")))

cmp <- ldply(cmp,
             .id = "effect",
             .fun = function(x) {
                 names(x)[2] <- "level"
                 return(x)
             })
cmp$level <- factor(cmp$level, levels = unique(cmp$level))
cmp
```

```{r echo = -c(1:2), fig.cap = cap}
cap <-
    "Mean proportion of fruits with 35% or less of the shell surface discolored as a function of the fungicide and hed level for two years."
cap <- fgn_("finalresult", cap)

resizePanels(
    useOuterStrips(
        segplot(level ~ lwr + upr | yr + effect,
                centers = fit,
                data = cmp,
                draw = FALSE,
                xlab = "Proportion of fruits with 35% or less damage",
                ylab = "Levels of the factors",
                scales = list(y = list(relation = "free"),
                              alternating = 1),
                ann = sprintf("%0.2f %s", cmp$fit, cmp$cld)
                )
    ), h = c(2, 4)) +
    layer(panel.text(x = centers[subscripts],
                     y = z[subscripts],
                     labels = ann[subscripts],
                     pos = 3))
```

****
# Session information

```{r, echo=FALSE, results="hold"}
cat(format(Sys.time(), format = "%A, %d de %B de %Y, %H:%M"),
    "----------------------------------------", sep = "\n")
sessionInfo()
```

<!------------------------------------------- -->
[Paulo S. F. Lichtemberg]: http://lattes.cnpq.br/8132272273348880
[Walmes M. Zeviani]: http://www.leg.ufpr.br/doku.php/pessoais:walmes
